Europe
========================================================

There are large discrepancies between reported import and export volumes.
Should we restric to a year? see France section below

```{r}
setwd("../..")
load("enddata/sawnwood_flows_Europe.RDATA")
library(plyr)
library(ggplot2)
library(reshape2)
```


# EU Trade Flows 
### from `r min(sawnwood_europe$year)` to `r max(sawnwood_europe$year)`
```{r results='asis'}
# Total volumes reported by EU countries to other non EU countries
tfeu2o <- ddply(subset(sawnwood_europe, reporting_EU & !partner_EU),
                .(item, element), summarise, value_EU = sum(value) )

# Total volumes reported by other countries - an EU country is their partner
tfo2eu <- ddply(subset(sawnwood_europe, !reporting_EU & partner_EU),
                .(item, element), summarise, value_partner = sum(value) )

tfo2eu$element <- sub("Export", "aaa", tfo2eu$element)
tfo2eu$element <- sub("Import", "Export", tfo2eu$element)
tfo2eu$element <- sub("aaa", "Import", tfo2eu$element)
tfeu <- mutate(merge(tfeu2o, tfo2eu), 
               diff_reporting_partner = value_EU - value_partner,
               diff_percent = diff_reporting_partner/value_EU*100) 
# Ther is certainly somehting wrong with these figures.
kable(tfeu)
```


# In 2011
```{r}

```

# Trade flows from France to EU countries
```{r}
# Total volumes reported by France to EU 
ddply(subset(sawnwood_europe, reporting_country=="France" & partner_EU &
                 year==2011),
      .(item, element), summarise, value_France = sum(value) )

# Total volumes reported by other EU countries - France is their partner
ddply(subset(sawnwood_europe, reporting_EU & partner_country=="France" & 
                 year==2011),
      .(item, element), summarise, value_partner = sum(value) )
```

### Trade flows from France to EU countries all years
How about trade flows with the rest of the world?
```{r fig.width=11}
# Put this into a function
# plot_tf_eu
# tfc2eu 
# tfeu2c

# Total volumes reported by France to EU 
tffreu <- ddply(subset(sawnwood_europe, reporting_country=="France" & partner_EU),
             .(item, element, year), summarise, value_France = sum(value) )

# Total volumes reported by other EU countries - France is their partner
tfeufr <- ddply(subset(sawnwood_europe, reporting_EU & partner_country=="France"),
      .(item, element, year), summarise, value_partner = sum(value) )

# Replace "Export" by "Import"
tfeufr$element <- sub("Export", "aaa", tfeufr$element)
tfeufr$element <- sub("Import", "Export", tfeufr$element)
tfeufr$element <- sub("aaa", "Import", tfeufr$element)
tffr <- merge(tffreu, tfeufr)

ggplot(data = tffr) +
    geom_point(aes(x = year, y = value_France, col="Reported by France")) + 
    geom_point(aes(x=year, y=value_partner, col="Reported by EU partners")) +
    facet_grid(element ~ item, scales="free_y")
```

### Table of Differences 
```{r results='asis'}
tffr <- mutate(tffr, diff_France_partner = value_France - value_partner)
# These could be written to Excel
kable(dcast(tffr[c("item", "element", "year", "diff_France_partner")],
                  item + element ~ year, value.var = "diff_France_partner"))
```
