__Metadata:__
* Author Paul Rougieux, European Forest Institute
* Data source: FAOSTAT
* Instructions to load data and libraries are indicated in the source file [sawnwood.Rmd](sawnwood.Rmd). 
* More information in the [readme](../../README.html)


```{r load_data, echo=FALSE}
setwd("../..")
load("rawdata/sawnwood_trade_us.RDATA")
```

```{r libraries, echo=FALSE}
library(ggplot2)
library(plyr)
library(xtable)
opts_chunk$set(fig.width=11, echo=FALSE)
```



```{r}
swdus <- sawnwood_trade_us
swdus$partner_country <- as.character(swdus$partner_country)
# remove "Others (adjustment)" and "unspecified" in the partner_country colunn
swdus <- subset(swdus, !partner_country %in% c("Others (adjustment)", "Unspecified"))
sorttrade <- function(dtf, year_, element_){
    arrange(subset(dtf, year == year_ & element == element_), -value)
    }
```


Major sawnwood trade partners of the `r unique(swdus$reporting_country)` 
============================================================

Tables
------

```{r, echo=FALSE}
# Total import in 2011
major <- function(dtf, n=6){
    dtf <- summarize(dtf, 
                   Year = year,
                   Country = partner_country, 
                   Percent = paste0(round(value/sum(dtf$value)*100, 2),"%"),
                   Volume_in_thousand_m3 = round(value/1e3))
    print(xtable(head(dtf, n), digits=0), type = "html")
    }
```

### In 2000
```{r, results='asis'}
major(sorttrade(swdus, 2000, "Import Quantity"))
major(sorttrade(swdus, 2000, "Export Quantity"))
```

### In `r max(swdus$year)`
```{r, results='asis'}
usimp <- sorttrade(swdus, max(swdus$year), "Import Quantity")
usexp <- sorttrade(swdus, max(swdus$year), "Export Quantity")
major(usimp)
major(usexp)
```


Plots
-----------
### Canada alone
```{r}
ggplot(data=subset(swdus, partner_country == "Canada")) + 
    aes(x=year, y=value/1e6, linetype=element) + geom_line() +
    ylab(expression(paste("Trade volume in million ",m^3)))
```

#### heatmap
```{r heatmapcanada, fig.height=2}
ggplot(data=subset(swdus, partner_country == "Canada")) + 
    aes(year, partner_country) +
    geom_tile(aes(fill = value/1e6), colour = "white") +
    facet_wrap(~element) +
    scale_fill_gradient(low = "white", high = "chocolate4", 
                        name = expression(paste("Trade volume in million ",m^3))) 
```

### Largest import + largest export partners except Canada
```{r }
big_imp_exp = c(head(usimp$partner_country), head(usexp$partner_country))
ggplot(data=subset(swdus, partner_country != "Canada" &
                   partner_country %in% big_imp_exp)) + 
    aes(x=year, y=value/1e6, color=partner_country, linetype=element) +
    geom_line() + facet_wrap(~ partner_country) + 
    ylab(expression(paste("Trade volume in million ",m^3)))
```

#### heatmap
```{r heatmap}
# All countries except canada
ggplot(data=subset(swdus, partner_country != "Canada" &
                   partner_country %in% big_imp_exp)) + 
    aes(year, partner_country) +
    geom_tile(aes(fill = value/1e6), colour = "white") +
    facet_wrap(~element) +
    scale_fill_gradient(low = "white", high = "chocolate4", 
                        name = expression(paste("Trade volume in million ",m^3))) 
```


### "Others (adjustment)" and "Unspecified" partner countries
```{r}
ggplot(data=subset(sawnwood_trade_us, 
                   partner_country %in% c("Others (adjustment)", "Unspecified"))) + 
    aes(x=year, y=value/1e6, color=partner_country, linetype=element) +
    geom_line() + facet_wrap(~ partner_country) + 
    ylab(expression(paste("Trade volume in million ",m^3)))
```


