

```{r echo=FALSE}
# This report template has to be called from knit_report.R
# product_trade_data comes from knit_report.R
library(ggplot2)
```

```{r functions, eval=TRUE}
ptd <- product_trade_data
# remove "Others (adjustment)" and "unspecified" in the partner_country colunn
ptd <- subset(ptd, !partner_country %in% c("Others (adjustment)", "Unspecified"))


sorttrade <- function(dtf, year_, element_){
    arrange(subset(dtf, year == year_ & element == element_), -value)
    }

# Total import in 2011
major <- function(dtf, n=6){
    dtf <- summarize(dtf, 
                   Year = year,
                   Country = partner_country, 
                   Percent = paste0(round(value/sum(dtf$value)*100, 2),"%"),
                   Volume_in_thousand_m3 = round(value/1e3))
    print(xtable(head(dtf, n), digits=0), type = "html")
    }

```


Major `r unique(ptd$item)` trade partners of `r unique(ptd$reporting_country)` 
========================================================


### 2000 Imports
```{r, results='asis'}
major(sorttrade(ptd, 2000, "Import Quantity"))
```

### 2000 Exports
```{r, results='asis'}
major(sorttrade(ptd, 2000, "Export Quantity"))
```


### `r max(ptd$year)` Imports
```{r, results='asis'}
last_imp <- sorttrade(ptd, max(ptd$year), "Import Quantity")
major(last_imp)
```

### `r max(ptd$year)` Exports
```{r, results='asis'}
last_exp <- sorttrade(ptd, max(ptd$year), "Export Quantity")
major(last_exp)
```


Plots
-----------
### Largest import + largest export partners 
This is based only on the last available year. 
May have to change this for the last 10 years 
for countries where trade partners change a lot.

```{r timelineMajor}
impo <- sorttrade(ptd, max(ptd$year), "Import Quantity")
expo <- sorttrade(ptd, max(ptd$year), "Export Quantity")
# large exporter and importer 
big_imp_exp = c(head(impo$partner_country), head(expo$partner_country))

ggplot(data=subset(ptd, partner_country %in% big_imp_exp)) + 
    aes(x=year, y=value/1e3, color=partner_country, linetype=element) +
    geom_line() + facet_wrap(~ partner_country) + 
     ylab(expression(paste("Trade volume in 1000",m^3)))
```

#### heatmap
```{r heatmap2}
# All countries except canada
ggplot(data=subset(ptd, partner_country %in% big_imp_exp)) + 
    aes(year, partner_country) +
    geom_tile(aes(fill = value/1e3), colour = "white") +
    facet_wrap(~element) +
    scale_fill_gradient(low = "white", high = "chocolate4", 
                        name = expression(paste("Trade volume in 1000 ",m^3))) 
```


### "Others (adjustment)" and "Unspecified" partner countries
```{r timelineAdjustment}
ggplot(data=subset(product_trade_data, 
                   partner_country %in% c("Others (adjustment)", "Unspecified"))) + 
    aes(x=year, y=value/1e3, color=partner_country, linetype=element) +
    geom_line() + facet_wrap(~ partner_country) + 
    ylab(expression(paste("Trade volume in 1000 ",m^3)))
```


